Assesing Health Care Coverage and Access Utilizing the National Health Interview Survey

### 2017-03-08 paper/paper.tex
update app and incude html of graphs

@@

@@

@@

A description of previous papers or projects related to your project.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  IPUMS Health Surveys. Retrieved February 27, 2017 from
  \url{https://ihis.ipums.org/ihis/index.shtml}
\item
  David Himmelstein,Steffie Woolhandler. January 23, 2017. Repealing the
  Affordable Care Act will kill more than 43,000 people annually.
  Retrieved March 6, 2017 from
  \url{https://www.washingtonpost.com/posteverything/wp/2017/01/23/repealing-the-affordable-care-act-will-kill-more-than-43000-people-annually/?utm_term}=.a2fbb24cd075
\end{enumerate}
\newpage
\singlespacing 
\end{document}svm

>>>>

To answer the question, \emph{``Does higher educational attainment cause
greater chance of acquiring medical insurance?''}, we filtered the
dataset, only looking a three specific variables, the \emph{highest
educational attainment}, \emph{health insurance coeverage}, and
\emph{health insurance offered at workplace}.

@@

\includegraphics{paper_files/figure-latex/healthcovereduattainment-1.pdf}
\textbf{Figure 3. Distribution of Health Coverage per Education Level}

\includegraphics{paper_files/figure-latex/workhealthinsuranceeduattainment-1.pdf}
\textbf{Figure 4. Educational Attainment vs.~Workplace-offered Health
Insurance}

@@

The first graph that we see here represents whether or not an individual
obtained health coverage based on their education level. Immediately, we
see that the number 2 subgraph contains a lot more results than the
number 1 subgraph, which indicates that there were more people who
received health coverage than did not. Under the number 1 subgraph, we
see the varying number of people who did not receive health coverage
given their level of education. There appears to be a notable peak at
the number 14 which is indicative of those who have completed up to a
high school diploma. A second notable peak is at the number 15 which
represents those who have attended college but had not received a
degree. Such might be the case because most people finish school up
until high school and also attend college, so it would make sense that
the bulk of the sample surveyed would reside in that educational
attainment categories. It would furthermore make sense that they did not
receive health coverage because nowadays, people obtain health/medical
insurance from their workplaces since getting these plans on an
individual basis is a lot more expensive. Those who did not finish
college or only obtained a high school diploma might have not been able
to get the best occupation due to their lack of education and skills in
the workforce, which affected their ability to get health coverage from
their employer.

In the number 2 subgraph, we see three notable peaks--one at the
x-coordinate values 14, 15, and 18. These peaks correspond to the bulk
of those who received health/medical insurance given that they either
graduated high school, went to college but did not get a degree, or got
a Bachelor's degree. What is interesting about this is that there were a
lot of people who did not AND did get health coverage given that they
graduated from high school or went to college but did not get a degree.
This might reveal to us that even with a high school degree it is
possible to get a steady job that includes medical/health insurance
coverage. However, analyzing the data further, it makes sense that a lot
of people who received health coverage received at least a Bachelor's
degree, because that level of education makes it a lot easier to get
jobs that pay well and include a lot of benefits such as medical
insurance.

The second graph we are looking at compares educational attainment with
whether or not a person received health insurance from their workplace.
The number 1 subgraph represents the count of people who did
\emph{receive health insurance at their workplace} given their highest
level of educational attainment, while number 2 subgraph represents the
count of people who \emph{did not receive health insurance at work}
given their highest level of educational attainment. Again, in both
graphs, x-coordinate values 14, 15, and 18, corresponding to having
graduated high school, attending college without getting a degree, and
getting a bachelor's degree respectively, are the most prominent peaks
in the graph. This is due to the fact that most people finish at least
up until high school, attend college at some point and realize that it
wasn't for them, or obtain their bachelor's degree. Very \emph{few}
people quit grade school in the middle of it and very few people get a
higher education such as a master's or doctoral degree, which is why the
sample count is so low at those other values. Based on the number 1
subgraph however, we see that the most people who got health insurance
from their workplace are those who graduated with a Bachelor's degree.
This reveals the importance of obtaining a Bachelor's degree when
getting a job. Other interesting points of analysis are that more high
school graduates received workplace-offered health insurance than those
who attended college but did not get a degree. One reasoning behind this
is that those who did not receive a degree from college might have been
those who quit in order to pursue their own businesses, and therefore
did not work for an existing corporation that would have otherwise
offered them health insurance. The number 2 subgraph shows that high
school graduates were the least likely to receive health insurance from
their workplace, which again points back to the fact that they probably
weren't able to get a great enough job that offered them benefits such
as health insurance. Most decent jobs nowadays require people to have at
least a bachelor's degree. The lowest of the three peaks, though, are
those who got a bachelor's degree, and it shows that with a higher level
of education, you will have a \emph{better} chance of getting a job with
better benefits.

ODDS RATIO: 1.19 / 3.66 = \textbf{0.33} The odds of getting coverage
being a high school grad is 0.33 times higher than the odds of getting
coverage having a bachelor's degree.

@@

Previous projects have been done that measured trends in health
insurance coverage over the course of time. The report was done by the
National Center for Health Statistics (NCHS), and presents the selected
estimates for health insurance coverage among the civilian
noninstitutionalized U.S. population based on data from 2015 National
Health Interview Survey (NHIS), as well as comparable estimates from the
NHIS 2010-2014 surveys. From their results, they were able to determine
that the trends displayed an overall increase in the population that had
health insurance coverage in comparison to previous years. They then
observe short term trends by age, poverty status, as well as race and
ethnicity to provide a deeper insight of health insurance coverages, and
what might be influencing access to them.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  IPUMS Health Surveys. Retrieved February 27, 2017 from
  \url{https://ihis.ipums.org/ihis/index.shtml}
\item
  David Himmelstein,Steffie Woolhandler. January 23, 2017. Repealing the
  Affordable Care Act will kill more than 43,000 people annually.
  Retrieved March 6, 2017 from
  \url{https://www.washingtonpost.com/posteverything/wp/2017/01/23/repealing-the-affordable-care-act-will-kill-more-than-43000-people-annually/?utm_term}=.a2fbb24cd075
\item
  Robin A. Cohen. May 2016. Health Insurance Coverage: Early Release of
  Estimates From the National Health Interview Survey, 2015. Retrieved
  March 2, 2017 from
  \url{https://www.cdc.gov/nchs/data/nhis/earlyrelease/insur201605.pdf}
\end{enumerate}
\newpage
\singlespacing 
\end{document}svm
